Adaptive group-regularized logistic elastic net regression

### 2019-03-06 docs/manuscript.Rnw
stability selection in parallel

	Examples of such a study with extra information available are investigated in
	Sections [! section refs here]. The first one is concerned with treatment 
	response prediction for colorectal cancer patients, based on sequenced
	microRNAs. In the second example we attempt to predict lymph node metastases,
	using RNAseq data. In both settings we have additional information on the 
	features in the form of a partitioning available. Furthermore, in both
	examples data is sparse, so we would like to 
	include the additional information to enhance predictive performance of our
	models. Inclusion of the partitions is not straightforward using standard 
	methods available in literature. At first sight, such
	an \textit{a priori} grouping of the features suggests the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which
	penalises and selects features at the group level. There are two limitations 
	to the group lasso: (a) it selects groups as a whole instead of single 
	features and (b) it is not adaptive: the amount of penalization a group
	receives depends on a `global' penalty parameter common to all groups.
	Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013}
	partly deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is,
	each group of variables receives its own penalty parameter. An apparent issue
	with this differential penalization is the estimation of the penalty 
	parameters. Naive estimation may be done by cross-validation. However, 
	cross-validation requires re-estimation of the model over a grid of penalty
	parameters. The size of this grid increases exponentially with the number of
	penalty parameters. Consequently, it quickly becomes computationally 
	infeasible. We therefore propose an efficient alternative: empirical Bayes
	estimation of the penalty parameters, which corresponds to hyperparameter
	estimation in the Bayesian prior framework. Because of the ubiquity of binary
	outcome data in omics research, we focus on the logistic elastic net. 
	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. We would like to emphasize that our
	method is different from recent developments in grouped penalized regression
	such as \cite{dondelinger_joint_nodate}, where the grouping is over the
	observations as opposed to the features

	In line with the above, our proposed method is (a) subjective: the few
	remaining tuning parameters may be cross-validated, (b) adaptive: prior
	information is automatically and subjectively weighed with respect to its
	informativeness, (c) fast compared to full Bayesian analysis or 
	cross-validation, and (d) requires minimal input by the user: only the data 
	and grouping of the features is required.

	We introduce the Bayesian logistic (generalised) elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} and Sections [! section ref here] and 
	[! section ref here], we demonstrate the approach in a
	simulation study and a two real data examples, respectively. In Section 
	\ref{sec:discussion} we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

	\section{Model}\label{sec:model}
	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
	In logistic regression the outcome variables are assumed to be binary or sums
	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
	the binomial distribution with number of trials $m$ and probability 
	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
	reduces to a binary logistic regression model. Throughout the rest of the 
	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	In elastic net regression, the penalised likelihood is maximised to yield
	parameter estimates:
	$$
	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
	\frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
	$$
	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$. The likelihood term is thus 
	complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the
	parameters, multiplied by respective penalty parameters $\lambda_1$ and
	$\lambda_2$. We have scaled the penalty parameters by $\frac{1}{2}$, to make
	them correspond with the Bayesian version in the following. The maximiser of
	the penalised likelihood corresponds to the posterior mode of a Bayesian
	elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
	\begin{equation}\label{eq:elasticnetprior}
	  \beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1
	  |\beta_j| + \lambda_2 \beta_j^2)\].
	\end{equation}
	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is 
	dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the 
	Supplementary Material (SM), along with further details on the elastic net
	prior). The elastic net combines $L_1$ and $L_2$-norm penalisation such that
	the model parameters are shrunken towards zero. The $L_1$-norm may set some of
	the estimates exactly to zero, thus automatically selecting features. The 
	$L_2$-norm ensures that correlated features behave similarly, as one would
	generally require. We extend the elastic net model to the generalised elastic
	net to allow for different penalty parameters per group of features.

	\subsection{The generalised elastic net}\label{sec:genen}
	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$. To apply differential penalisation we include a 
	group-specific weight $w_g \in \mathbb{R}_{>0}$. The model parameters are now
	estimated by:
	\begin{align}
	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
	  \label{eq:weightedlik}\\
	  & = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	  \nonumber
	\end{align}
	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific
	weights may be interpreted as penalty multipliers. Throughout the following we
	assume that the geometric mean of the multipliers, weighted by their 
	respective group sizes, is one, such that the average shrinkage of the model
	parameters is determined by the `global' $\lambda_1$ and $\lambda_2$. That is,
	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
	$L_1$-norm term to ensure that penalisation on the parameter level is the same
	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
	(\ref{eq:weightedlik}).

	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see Section 3 of the SM for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
	respectively.

@@

	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
	Variational Bayes was shown to underestimate the posterior variance of the 
	parameters, both numerically and theoretically, in several settings 
	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
	wang_inadequacy_2005}. This coincides with our experience that the global 
	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
	prevent overestimation we use the parametrisation of the elastic net in 
	\cite{friedman_regularization_2010} as introduced in Section 
	\ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
	cross-validation of the regular elastic net model, such that the overall 
	penalisation is determined by cross-validation of only $\lambda$. By combining
	cross-validation of the global penalty parameter $\lambda$ with empirical 
	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
	robust to underestimation of the variational posterior variances. The 
	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
	choices of $\alpha$. We recommend the latter. Otherwise we note that, in our 
	experience, the choice $\alpha=0.5$ often gives good results.

	For estimating the penalty multipliers, the intractable posterior expectation
	in (\ref{eq:ebEM}) is approximated by the variational posterior:
	$$
	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \]
	= \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
	\lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	$$
	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full 
	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
	of the new penalty multipliers is now given by:
	\begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}
	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
	closed form, this convex problem is easily solved by a numerical optimisation 
	routine. The full procedure is summarized in Section 8 of the SM.

	\subsection{Feature selection}
	Feature selection is often desirable in high-dimensional prediction problems 
	and omics data problems are no exception. A selection of biomarkers may lead 
	to a large cost reduction by supporting targeted assays. Bayesian feature
	selection is often done by inspection of posterior credible intervals. 
	Similarly as in frequentist hypothesis testing, we may select a feature if 
	zero is not contained in the credible interval. However, the Bayesian lasso's
	credible intervals (a special case of the elastic net) are known to suffer 
	from low frequentist coverage in sparse settings 
	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
	omics research, so selection by credible intervals is generally inconsistent. 
	We therefore propose to select features in the frequentist paradigm.

@@

	\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
  The presented data is from a deep sequencing analysis on microRNAs 
  \cite[]{neerincx_combination_2018}. The study was done in 88 treated 
  colorectal cancer patients, with the aim of predicting treatment response,
  coded as either non-progressive/remission (70 patients) or progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. In addition to the 2114 microRNAs, four clinical unpenalized 
  covariates were included in the analysis: prior use of adjuvant therapy 
  (binary), the type of systemic treatment regimen (ternary), age, and primary 
  tumor differentiation (binary).

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, and to the best performing
  group lasso extension, the composite minimax
  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. 
  For the methods with an extra tuning parameter 
  $\alpha$, we picked the one (or two) that gave the best performance and refer
  the reader to SM Section [! section ref here] for more comparisons to 
  different $\alpha$s and to the sparse
  group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and the group
  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.

  To estimate performance, we split the data into 61 training, and 27 test 
  instances stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present AUC for a range 
  of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. More 
  performance measures comparisons are presented in Section [! section ref here]
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation using this data set in SM Section 
  [! section ref here]. A small sidenote: We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global penalty parameter on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty. 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes. This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. We noticed this is in general
  for the group-lasso and its extensions (see SM [! section ref here]). Overall,
  it seems as though the inclusion of the extra information on the features 
  benefited predictive performance: both \texttt{GRridge} and \texttt{gren} 
  outperform their respective non-group-regularized versions, albeit it only
  slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@

  \section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
  In this application, the aim is to predict lymph node metastasis in patients 
  with HPV-negative oral cancer using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  normalised TCGA RNASeqv2 profiles of head-and-neck squamous cell 
  carcinomas on 133 patients with HPV-negative tumours in the oral cavity. Of 
  these patients, 76 suffered from lymph node metastasis, while 57 did not. 
  For a thorough introduction of these data, see \cite{te_beest_improved_2017}.

  As co-data we considered two sources: (a) the cis-correlation between the
  the RNASeqv2 data and DNA copy numbers on the same patients, 
  quantified by Kendall's $\tau$ and binned into five groups, and (b) 
  $p$-values from an external microarray data set described in 
  \cite{mes_prognostic_2017}, again binned into five groups. We expect features 
  with a large positive Kendall's $\tau$ to be more important in 
  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
  features with low $p$-values to be more important.

>>>>

	Two such studies are introduced in Sections \ref{sec:colorectal} and
	\ref{sec:lymphnode}. The first study is concerned with treatment 
	response prediction for colorectal cancer patients, based on sequenced
	microRNAs. In the second study we attempt to predict lymph node metastases,
	using RNAseq data. In both settings additional information on the 
	features is available in the form of a partitioning. Furthermore, in both
	examples data is sparse, so any extra included information may help to boost
	predictive performance. Inclusion of such feature partitions in 
	the analysis is not straightforward with standard methods.

	At first sight, such an \textit{a priori} grouping of the features suggests 
	the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
	the group lasso penalises and selects features at the group level. 
	There are two limitations to this approach: the group lasso (a) selects groups
	as a whole instead of single 
	features and (b) does not penalise adaptively: all groups are penalised 
	equally. Extensions such as the sparse group lasso 
	\cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is,
	each group of variables receives its own penalty parameter. An apparent issue
	with this differential penalization is the estimation of the penalty 
	parameters. Naive estimation may be done by cross-validation. However, 
	cross-validation requires re-estimation of the model over a grid of penalty
	parameters. The size of this grid increases exponentially with the number of
	penalty parameters. Consequently, it quickly becomes computationally 
	infeasible. We therefore propose an efficient alternative: empirical Bayes
	estimation of the penalty parameters, which corresponds to hyperparameter
	estimation in the Bayesian prior framework. Because of the ubiquity of binary
	outcome data in omics research, we focus on the logistic elastic net.

	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but just as VBVS, these methods lack 
	adaptive inclusion of external information on the features. We emphasize that 
	our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
	features, the observations are partitioned.

	In line with the above, our proposed method is (a) subjective: the 
	hyper-parameters are estimated from the data and not chosen on objective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required.

	The rest of the paper is structured as follows. We introduce the Bayesian 
	logistic generalised elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} we compare the method in a simulation study. In Sections
	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
	the two previously introduced data sets. In Section \ref{sec:discussion} 
	we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

	\section{Model}\label{sec:model}
	In logistic regression the outcome variables are assumed to be binary or sums
	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
	the binomial distribution with number of trials $m$ and probability 
	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
	reduces to a binary logistic regression model. Throughout the rest of the 
	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
	group-specific weight. In generalised elastic net regression, the penalised 
	likelihood is maximised to yield parameter estimates:
	\begin{align}
	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
	  \label{eq:weightedlik}\\
	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	  \nonumber
	\end{align}
	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
	observed data $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
	group-specific weights may be interpreted as penalty multipliers. Note that
	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.

  Throughout the following we assume that the geometric mean of the multipliers, 
	weighted by their 
	respective group sizes, is one, such that the average shrinkage of the model
	parameters is determined by $\lambda_1$ and $\lambda_2$. That is,
	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
	$L_1$-norm term to ensure that penalisation on the parameter level is the same
	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
	(\ref{eq:weightedlik}). The elastic net combines $L_1$ and $L_2$-norm 
	penalisation such that the model parameters are shrunken towards zero. 
	The $L_1$-norm may set some of the estimates exactly to zero, thus 
	automatically selecting features. The $L_2$-norm ensures that correlated 
	features behave similarly, as one would generally require.

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see SM Section [! section reference here] for details on this prior) may be 
	written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see Section [! section ref here] 
	of the SM for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
	respectively.

@@

	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
	Variational Bayes was shown to underestimate the posterior variance of the 
	parameters, both numerically and theoretically, in several settings 
	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
	wang_inadequacy_2005}. This coincides with our experience that the global 
	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
	prevent overestimation we use the parametrisation of the elastic net in 
	\cite{friedman_regularization_2010} as introduced in Section 
	\ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
	cross-validation of the regular elastic net model, such that the overall 
	penalisation is determined by cross-validation of only $\lambda$. By combining
	cross-validation of the global penalty parameter $\lambda$ with empirical 
	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
	robust to underestimation of the variational posterior variances. The 
	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
	choices of $\alpha$. We recommend the latter.

	To estimate the penalty multipliers, the intractable posterior expectation
	in (\ref{eq:ebEM}) is approximated using the variational posterior:
	$$
	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \]
	= \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
	\lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	$$
	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full 
	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
	of the new penalty multipliers is now given by:
	\begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}
	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
	closed form, this convex problem is easily solved by a numerical optimisation 
	routine. The full procedure is summarized in Section 8 of the SM.

	\subsection{Feature selection}
	Feature selection is often desirable in high-dimensional prediction problems 
	and omics data problems are no exception. For example, a selection of 
	biomarkers may lead 
	to a large cost reduction by supporting targeted assays. Bayesian feature
	selection is often done by inspection of posterior credible intervals. 
	Similarly as in frequentist hypothesis testing, we may select a feature if 
	zero is not contained in the credible interval. However, the Bayesian lasso's
	credible intervals (a special case of the elastic net) are known to suffer 
	from low frequentist coverage in sparse settings 
	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
	omics research, so selection by credible intervals is generally inconsistent. 
	We therefore propose to select features in the frequentist paradigm.

@@

	\section{Application to microRNAs in colorectal cancer}
	\label{sec:colorectal}
  As shortly touched upon in Section \ref{sec:introduction}, the data is from a 
  deep sequencing analysis on microRNAs 
  \cite[]{neerincx_combination_2018}. This study in 88 treated 
  colorectal cancer patients was conducted with the aim of predicting treatment 
  response, coded as either non-progressive/remission (70 patients) or 
  progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. In addition to the 2114 microRNAs, four clinical unpenalized 
  covariates were included in the analysis: prior use of adjuvant therapy 
  (binary), the type of systemic treatment regimen (ternary), age, and primary 
  tumor differentiation (binary).

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, and to the best performing
  group lasso extension, the composite minimax
  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. 
  For the methods with an extra tuning parameter 
  $\alpha$, we picked the one (or two) that gave the best performance and refer
  the reader to SM Section [! section ref here] for more comparisons with 
  different $\alpha$s and comparisons to the sparse
  group lasso by \cite{simon_sparse-group_2013} and group
  exponential lasso \cite[]{breheny_group_2015}.

  To estimate performance, we split the data into 61 training and 27 test 
  instances, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present AUC for a range 
  of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. More 
  performance measures are presented in Section [! section ref here]
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation using in SM Section 
  [! section ref here]. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global penalty parameter on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. We noticed that this is in general the 
  case for the group-lasso and its extensions (see SM [! section ref here]). 
  Overall, it seems as though the inclusion of the extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@

  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
  The aim of this study is to predict lymph node metastasis in patients 
  with HPV-negative oral cancer using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
  did not. For a thorough introduction of these data, see 
  \cite{te_beest_improved_2017}.
